========================================================
# **Peer Assessment 1**, *Reproducible Research*
# *Data Science Specialization* at *Coursera*
## ======================================================
## I. Load and Preprocess Data 
### 1 and 2
```{r read, echo=TRUE}
dat <- read.csv("activity.csv", quote="", header=T)
names(dat)<- c("Steps_in_5_min_interval", "Measurement_Date", "Interval_minute")
dim(dat)
head(dat, n=3)
a <- aggregate(as.numeric(dat[,1])~dat[,2], data=dat,sum)
names(a) <- c("days", "steps")
dim(a)
head(a, n=3)
```

## ======================================================
## II. What is mean total number of steps taken per day?  

### 1. Here is a histogram of number of steps per day with breaks at 1000 days for clarity. 

```{r histogram, fig.width=7, fig.height=6}
hist(a[,2], main = "STEPS PER DAY", breaks = seq(0,25000,1000),
     xlab = "Per 1000 days", ylab ="Number of steps",
     col="blue")
```
```{r mean, echo=TRUE}
m <- format(mean(a$steps), digits=2, nsmall=2)
me <- format(median(a$steps), digits=2)
```
### 2. On an average, `r m` steps and a  median of `r me` steps were taken per day.

## ======================================================
## III. What is the average daily activity pattern? 
### 1.Time series plot
```{r timeseries, fig.width=7, fig.height=6}
b <- aggregate(as.numeric(dat[,1])~dat[,3], data=dat, mean)
head(b, n=3)
dim(b)
plot(b[,2]~b[,1], type="l", main="AVERAGE DAILY ACTIVITY PATTERN",
     xlab = "5-min Time interval",
     ylab = "Average daily steps")
```
```{r max, echo=FALSE}
w <- which.max(b[,2])
```
### 2. The `r b[w,1]`-minute interval, on average across all the days in the dataset, contains the maximum number of steps.
## ======================================================
## IV. Imputing missing values
### 1. Calculate missing values
```{r missing, echo=TRUE}
miss <- dat[is.na(dat),]
missr <- nrow(miss)
```
### There are `r missr` rows with missing values in the dataset.  
### 2 and 3. Impute missing values, make new dataset with imputed values
```{r imputemean, echo=TRUE}
# compute mean steps for 5 minute intervals
agg <- aggregate(as.numeric(dat[,1])~dat[,3], data=dat, mean)
# match interval minutes, replace NA with corresponding mean steps
r <- replace(miss[,1], miss$Interval_minute %in% agg[,1], agg[,2])
miss$Steps_in_5_min_interval <- r
# make new dataset with replaced values
compdat <- dat[complete.cases(dat),]
all <- rbind(miss, compdat)
all <- all[order(all$Measurement_Date),]
```
### 4. Make histogram of imputed data
```{r hist2,  fig.width=7, fig.height=6 }
all_a <- aggregate(as.numeric(dat[,1])~dat[,2], data=dat,sum)
names(all_a) <- c("days", "steps")
hist(all_a[,2], main = "STEPS PER DAY", breaks = seq(0,25000,1000),
     xlab = "Per 1000 days", ylab ="Number of steps",
     col="blue")
```
```{r silent, echo=FALSE}
m1 <- format(mean(all_a$steps), digits=2, nsmall=2)
me1 <- format(median(all_a$steps), digits=2)
```
### The `r m1` mean steps  and  `r me1` median steps appear unchanged after imputing.
## ======================================================
## V. Are there differences in activity patterns between weekdays and weekends? 
### 1. Create new variable with two labels "weekend"" and "weekdays"
```{r days, echo=TRUE}
#Convert Measurment_Date variable to date format, label as weekday vs weekend
dates0 <- all$Measurement_Date
dates <- as.POSIXct(strptime(as.character((dates0)), "\"%Y-%m-%d\""))
w <- weekdays(dates)
new <- ifelse(w %in% c("Saturday","Sunday"),"weekend", "weekday")
allb <- cbind(all,new)
```
### 2. Make time series plot for weekend vs weekdays
```{r timeseries2, fig.width=12, fig.height=6}
allbwy <- allb[allb$new == "weekday",]
allbwend<- allb[allb$new == "weekend",]
aggweekday <- aggregate(as.numeric(allbwy[,1])~allbwy[,3], data=allbwy,mean)
aggweekend <- aggregate(as.numeric(allbwend[,1])~allbwend[,3], data=allbwend,mean)
aggweekday <- cbind(aggweekday, allbwy$new)
aggweekend <- cbind(aggweekend, allbwend$new)
names(aggweekday) <- c("Interval", "Number of steps", "day")
names(aggweekend) <- c("Interval", "Number of steps", "day")
tseries <- rbind(aggweekday, aggweekend)
library(ggplot2)
g <- ggplot(tseries,aes(tseries[,1], tseries[,2])) 
g + facet_grid(.~day) + geom_line(aes(colour=day)) + labs(x="Interval") + labs(y="Number of Steps")
```